---
title: "Data preparation"
author: "Simon Rolph"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an example of data preparation of the input files. It reads files from the `data` directory and saves outputs to the `inputs` directory.

There is a directory called `data` which I use for holding downloaded files which I then process into `inputs`.

```{r}
library(sf)

```

## Species list

No processing is required to generate the species list. It's simply a case of writing a text file `inputs/species_list/sp_list.txt` and adding new lines with taxon keys eg.

```
5334220
5410907
8211070

```

## Shapefiles

```{r}
# read in the shapefile for the boundary that I downloaded 
AOI_boundary_file_location = "data/raw/boundaries/SG_CairngormsNationalPark_2010/SG_CairngormsNationalPark_2010.shp"

#read in the park boundary
boundary <- st_read(AOI_boundary_file_location)

#this prints the bounding box which can be used in the GEE script
boundary %>% st_bbox()

#export boundary as reprojected shapefile
boundary_buffered <- boundary %>%
  st_transform(4326) %>% #transform to wgs 1984
  st_buffer(10000)


st_write(boundary_buffered,"inputs/regions/AOI_fit/AOI_fit.shp",append=FALSE)
st_write(boundary_buffered,"inputs/regions/AOI_predict/AOI_predict.shp",append=FALSE)

#appriximate WKT of park with a bit of buffering
boundary_wkt <- boundary_buffered %>% st_simplify(dTolerance=1000) %>% # simplify
  st_geometry() %>% #only select the geometry to that st_as_text() works properly
  st_as_text()

#write the boundary of the AOI (area of interest) as WKT which can be used for querying GBIF
writeLines(boundary_wkt,"inputs/regions/AOI_fit/AOI_WKT.txt")
writeLines(boundary_wkt,"inputs/regions/AOI_predict/AOI_WKT.txt")

```

## Environmental data

Load layers which have been downloaded from google earth engine and reproject to OSGB. 

```{r}
#combine into one spatRaster
env_layers <- rast(list.files("data/raw/environmental/",full.names = T,pattern=".tif"))

#proximity measures
env_layers_proximity <- env_layers[[c('water','grass','flooded_vegetation','crops','shrub_and_scrub','built','bare','snow_and_ice')]]
names(env_layers_proximity) <- paste0("prox_",names(env_layers_proximity))
env_layers_proximity <- env_layers_proximity %>% aggregate(fact=9)
#create a Gaussian window to use as smoothing
w <- rayimage::generate_2d_gaussian(
  sd = 4,
  dim = c(11, 11),
  rescale_unity = T
)
image(w)

#apply the smoothing
#do a coarse moving window with max
env_layers_proximity2 <-  focal(env_layers_proximity,w=w, fun=max)

#then do a nicer smooth with mean
env_layers_proximity2 <-  focal(env_layers_proximity2,w=w, fun=mean)

# resample back to original resolution
env_layers_proximity2 <- resample(env_layers_proximity2,env_layers)

#add it to the other layers
env_layers <- c(env_layers,env_layers_proximity2)

# set cells with water are set to NA on all layers because we are only modelling terrestrial species
env_layers[env_layers$water>15] <- NA

# remove the water layer
env_layers <- env_layers[[names(env_layers)!="water"]]

#(optionally) aggregate to make coarser for quicker model runs (for testing purposes)
env_layers <- env_layers %>% aggregate(fact=4)

env_layers

#save the raster  
writeRaster(env_layers,filename = "inputs/environmental/env-layers.tif",overwrite=T)

```

