---
title: "SDM evaluation"
date: "`r Sys.Date()`"
output: html_document
params:
  species_metadata: species_metadata
  sdm_data: sdm_data
  full_model:  full_model
  sdm_pred:  sdm_pred
  bs_models:  bs_models
  sdm_var: sdm_var
---

```{r,echo=F,include=F}
knitr::opts_chunk$set(echo = TRUE)
species_metadata <- params$species_metadata
sdmdata <- params$sdm_data
full_model <- params$full_model
sdm_pred <- params$sdm_pred
bs_models <- params$bs_models
sdm_var <- params$sdm_var


library(terra)
library(leaflet)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ggeffects)
```

# `r species_metadata$species` 

`r species_metadata$vernacularName`

## Maps

### Species distribution

Where does the model predict the species to occur? Blue points indicate true presences.

```{r,echo=F, out.width = '100%'}
plet(rast(sdm_pred),tiles="Streets") %>% 
  addCircleMarkers(~x, ~y, data = sdmdata[sdmdata$pres==1,],stroke = FALSE, fillOpacity = 0.5,radius=3)
```

### Model variability

Where are the model outputs most variable? Blue points indicate true presences.

```{r,echo=F, out.width = '100%'}
plet(rast(sdm_var),tiles="Streets") %>% 
  addCircleMarkers(~x, ~y, data = sdmdata[sdmdata$pres==1,],stroke = FALSE, fillOpacity = 0.5,radius=3)
```

## Presence vs background plots

Blue density indicates environmental conditions of presence records. Grey indicates environment of background samples.

```{r,echo=F, out.width = '100%',warning=FALSE}
sdmdata2 <- sdmdata %>% select(-x,-y,-ID)
sdmdata2 <-sdmdata2 %>% pivot_longer(cols = c(-pres,-fold))


ggplot(sdmdata2 %>% filter(pres==1),aes(x = value)) +
  geom_density(data = sdmdata2 %>% filter(pres==0),fill = "black",alpha = 0.1) +
  geom_density(data = sdmdata2 %>% filter(pres==1),fill="blue",alpha= 0.2) +
  facet_wrap(~name,scales = "free")+
  theme_minimal()+
  theme(text = element_text(size = 5))
```

## Model summary

```{r}
summary(full_model)

plot(full_model)
```

## K fold predictons

For each fold, models are fitted with (k-1)/k of the data. The model then predicts the probability of occurrence of the excluded true presences. 

```{r,echo=F}
prediction_compare <- data.frame(fold =0,prediction = 0)[-1,]
for (i in 1:length(bs_models)){
  predictions <- predict(bs_models[[i]],newdata = sdmdata[sdmdata$fold == i & sdmdata$pres == 1,],type="response") 
  
  prediction_compare <- bind_rows(prediction_compare,data.frame(fold =i,prediction = predictions))
}

ggplot(prediction_compare,aes(x = fold,y = prediction,group = fold))+
  geom_boxplot(fill = "blue",alpha = 0.2)+
  theme_minimal()

```

## Marginal effects

```{r, results='asis',echo=F}
terms <- names(full_model$coefficients)
for (i in 2:length(terms)){
  margins <- ggpredict(full_model, terms=paste0(terms[i]," [all]"))
  print(plot(margins))
}

```

## Species meta data

```{r}
str(species_metadata)
```






